---
title: "Introduction to gfaR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gfaR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gfaR)
```

## Overview

`gfaR` is an R package for finding non-B DNA motifs in genomic sequences. It is a refactored version of the Genomic Feature Analyzer (GFA) C-program.

Regina Z. Cer, Duncan E. Donohue, Uma S. Mudunuri, Nuri A. Temiz, Michael A. Loss, Nathan J. Starner, Goran N. Halusa, Natalia Volfovsky, Ming Yi, Brian T. Luke, Albino Bacolla, Jack R. Collins and Robert M. Stephens.

Nucl. Acids Res. (2013) 41 (D1): D94-D100. doi: 10.1093/nar/gks955 

## Supported Motifs

The package can detect the following motif types:

- **A-Phased Repeats (APR)**: Associated with bent DNA.
- **Direct Repeats (DR)**: Associated with slipped DNA.
- **G-quadruplexes (GQ)**: Associated with four-stranded structures.
- **Inverted Repeats (IR)**: Associated with cruciforms.
- **Mirror Repeats (MR)**: Associated with triplexes.
- **Short Tandem Repeats (STR)**: Short repeating units.
- **Z-DNA**: Left-handed DNA helices.

## Usage

To analyze a sequence, use the `gfa_analyze` function:

```{r example}
sequence <- "AGTGCAACCCAGAGGGCAGGATTTCCTGCTGGACTTTGAAATCCAACCCGGTCACCTACCCGCGCGACTG"
results <- gfa_analyze(sequence)

# View detected G-quadruplexes
print(results$gq)
```

You can also call individual finding functions:

```{r individual}
apr_results <- find_apr(sequence)
print(apr_results)
```

## References

- Non-B DB v2.0: a database of predicted non-B DNA-forming motifs and its associated tools. Nucleic Acids Res. 2016.

Regina Z. Cer, Duncan E. Donohue, Uma S. Mudunuri, Nuri A. Temiz, Michael A. Loss, Nathan J. Starner, Goran N. Halusa, Natalia Volfovsky, Ming Yi, Brian T. Luke, Albino Bacolla, Jack R. Collins and Robert M. Stephens.

Nucl. Acids Res. (2013) 41 (D1): D94-D100. doi: 10.1093/nar/gks955 