---
title: "Introduction to gfaR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gfaR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gfaR)
```

## Overview

`gfaR` is an R package for finding non-B DNA motifs in genomic sequences. It is a refactored version of the Genomic Feature Analyzer (GFA) C-program.

Regina Z. Cer, Duncan E. Donohue, Uma S. Mudunuri, Nuri A. Temiz, Michael A. Loss, Nathan J. Starner, Goran N. Halusa, Natalia Volfovsky, Ming Yi, Brian T. Luke, Albino Bacolla, Jack R. Collins and Robert M. Stephens.

Nucl. Acids Res. (2013) 41 (D1): D94-D100. doi: 10.1093/nar/gks955 

## Supported Motifs

The package can detect the following motif types:

- **A-Phased Repeats (APR)**: Associated with bent DNA.
- **Direct Repeats (DR)**: Associated with slipped DNA.
- **G-quadruplexes (GQ)**: Associated with four-stranded structures.
- **Inverted Repeats (IR)**: Associated with cruciforms.
- **Mirror Repeats (MR)**: Associated with triplexes.
- **Short Tandem Repeats (STR)**: Short repeating units.
- **Z-DNA**: Left-handed DNA helices.

## Usage

To analyze a sequence, use the `gfa_analyze` function:

```{r example}
sequence <- "AGTGCAACCCAGAGGGCAGGATTTCCTGCTGGACTTTGAAATCCAACCCGGTCACCTACCCGCGCGACTG"
results <- gfa_analyze(sequence)

# View detected G-quadruplexes
print(results$gq)
```

### Using Included Test Data

The package includes sample data for testing and demonstration purposes. You can locate these files using `system.file()`:

```{r extdata}
# Get the path to the included test FASTA file
fasta_path <- system.file("extdata", "gfa_test.fasta", package = "gfaR")

# Read the sequence (simple helper for this example)
read_fasta <- function(file) {
    lines <- readLines(file)
    seq_lines <- lines[!grepl("^>", lines)]
    paste0(seq_lines, collapse = "")
}

test_seq <- read_fasta(fasta_path)
results <- gfa_analyze(test_seq)

# View the first few APR results
head(results$apr)
```

You can also find the original "gold standard" GFF and TSV outputs in the same directory to validate your results.

## References

- Non-B DB v2.0: a database of predicted non-B DNA-forming motifs and its associated tools. Nucleic Acids Res. 2016.

Regina Z. Cer, Duncan E. Donohue, Uma S. Mudunuri, Nuri A. Temiz, Michael A. Loss, Nathan J. Starner, Goran N. Halusa, Natalia Volfovsky, Ming Yi, Brian T. Luke, Albino Bacolla, Jack R. Collins and Robert M. Stephens.

Nucl. Acids Res. (2013) 41 (D1): D94-D100. doi: 10.1093/nar/gks955 