---
title: "Introduction to gfaR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gfaR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gfaR)
```

## Overview

`gfaR` is an R package for finding non-B DNA motifs in genomic sequences. It is a refactored version of the Genomic Feature Analyzer (GFA) C-program.

## Supported Motifs

The package can detect the following motif types:

- **A-Phased Repeats (APR)**: Associated with bent DNA.
- **Direct Repeats (DR)**: Associated with slipped DNA.
- **G-quadruplexes (GQ)**: Associated with four-stranded structures.
- **Inverted Repeats (IR)**: Associated with cruciforms.
- **Mirror Repeats (MR)**: Associated with triplexes.
- **Short Tandem Repeats (STR)**: Short repeating units.
- **Z-DNA**: Left-handed DNA helices.

## Usage

To analyze a sequence, use the `gfa_analyze` function:

```{r example}
sequence <- "AGTGCAACCCAGAGGGCAGGATTTCCTGCTGGACTTTGAAATCCAACCCGGTCACCTACCCGCGCGACTG"
results <- gfa_analyze(sequence)

# View detected G-quadruplexes
print(results$gq)
```

You can also call individual finding functions:

```{r individual}
apr_results <- find_apr(sequence)
print(apr_results)
```

## References

- Non-B DB v2.0: a database of predicted non-B DNA-forming motifs and its associated tools. Nucleic Acids Res. 2016.
